---
title: "Homework_10"
author: "Grace Durant"
date: "3/30/2022"
output: html_document
---

# Homework 10: For Loops and Randomization tests 
### Question 1 Directions: Using a for loop, write a function to calculate the number of zeroes in a numeric vector. Before entering the loop, set up a counter variable counter <- 0. Inside the loop, add 1 to counter each time you have a zero in the matrix. Finally, use return(counter) for the output.

```{r}
counter <- 0
num_vector<-c(0,0,0,0,3,4,5,3,6,7,5,3,2,4,2,1,4,6,9,0,3,4,7,8,0,2,34,56,6,9,0,2,3,56,7,8,9,0,2,2,4,4,6,7,9,9,0,0,8,7,5)
for (i in 1:length(num_vector)) {
  if(num_vector[i]==0) {
    counter <- counter+1
  }
}
print(counter)
```

### Question 2 Directions: Use subsetting instead of a loop to rewrite the function as a single line of code.

```{r}
num_vector_sub <-length(num_vector[num_vector==0])
print(num_vector_sub)
```

### Question 3 Directions: Write a function that takes as input two integers representing the number of rows and columns in a matrix. The output is a matrix of these dimensions in which each element is the product of the row number x the column number.

```{r}
#############################################
# FUNCTION: my_function
# input:two integers representing the number of rows and columns in a matrix
# output: a matrix of these dimensions in which each element is the product of the row number x the column number
# -------------------------------------------
my_function <- function(x=5, y=5) {
  mat <- matrix(sample(runif(10), size=x*y, replace=TRUE),
              nrow=x,
              ncol=y)
    for(i in 1:x) {
      for(j in 1:y){
        mat[i,j] <- i*j
      }
    } 
  return(mat)
}  
print(my_function())
```

### Question 4a-d Directions: 
* Simulate a dataset with 3 groups of data, each group drawn from a distribution with a different mean. The final data frame should have 1 column for group and 1 column for the response variable.
* Write a custom function that 1) reshuffles the response variable, and 2) calculates the mean of each group in the reshuffled data. Store the means in a vector of length 3.
* Use a for loop to repeat the function in b 100 times. Store the results in a data frame that has 1 column indicating the replicate number and 1 column for each new group mean, for a total of 4 columns.
* Use qplot() to create a histogram of the means for each reshuffled group. Or, if you want a challenge, use ggplot() to overlay all 3 histograms in the same figure. How do the distributions of reshuffled means compare to the original means?

```{r}
# Part a, simulating a data set 
three_groups <- c(rep("group_1",4),rep("group_2",5), rep("group_3",3))
print(three_groups)
response_variable <- c(runif(4)+3, runif(5)+7, runif(3)-0.5)
print(response_variable)
df<- data.frame(group=three_groups,res=response_variable)
print(df)

# part b, custom function
df_sim <- df
df_sim$res <- sample(df_sim$res)
print(df_sim)

sim <- tapply(df_sim$res,df$group,mean)
print(sim)
sim_vector <- as.vector(sim)
print(sim_vector)

# Part c, for loop
##################################################
# function: shuffle_data
# Purpose: use for loop to repeat function 100 times
# input: means of shuffled response variables in vector form
# output: 4-column data frame (replicate #, group mean 1, group mean2, group mean 3)
#------------------------------------------------- 
shuffle_data <-  function(data=NULL) {
  data$res <- sample(data$res)
                 
return(data)
}
shuffle_data(df) 

get_metric <- function(data=NULL) {
. <- summary(aov(data=data,res~group))
p_val <- .[[1]][["Pr(>F)"]][1]

return(p_val)
}
get_pval <- function(z=NULL) {
                    if(is.null(z)){
                      z <- list(x_obs=runif(1),x_sim=runif(1000))}
                      p_lower <- mean(z[[2]]<=z[[1]])
                      p_upper <- mean(z[[2]]>=z[[1]])
return(c(pl=p_lower,pu=p_upper))
                    }          
n_sim <- 100
x_sim <- rep(NA,n_sim) # vector of simulated slopes
df <- df
x_obs <- get_metric(df)
for (i in seq_len(n_sim)) {
x_sim[i] <- get_metric(shuffle_data(df))
}
head(x_sim)
p_value <- list(x_obs,x_sim)
get_pval(p_value)
library(ggplot2)
plot_ran_test <- function(z=NULL) {
                if(is.null(z)){
                  z <- list(rnorm(1),rnorm(1000)) }
df <- data.frame(ID=seq_along(z[[2]]),sim_x=z[[2]])

p1 <- ggplot(data=df) + 
       aes(x=sim_x)

p1 + geom_histogram(aes(fill=I("goldenrod"),
                        color=I("black"))) +
     geom_vline(aes(xintercept=z[[1]],
               col="blue")) 

                  }

plot_ran_test(p_value)


```
