---
title: "Homework_07"
author: "Grace Durant"
date: "3/2/2022"
output: html_document
---
# Homework 7: Data Manipulation using the dplyr  package 
```{r, echo=FALSE, message=FALSE}
library(tidyverse)
```
# Question 1
```{r}
data(iris)
class(iris)
```
## 150 Observations and 5 Variables 

# Question 2
```{r}
iris1 <- filter(iris, Species %in% c("virginica","versicolor"), Sepal.Length>6, Sepal.Width>2.5)
```
## 56 Observations and 5 variables 

# Question 3
```{r}
iris2<-select(iris1, Species, Sepal.Length, Sepal.Width)
```
## 56 Observations and 3 Variables

# Question 4
```{r}
iris3 <- arrange(iris2,by=desc(Sepal.Length))
 head(iris3)         
```
## 56 Observations and 3 Variables

# Question 5
```{r}
iris4<- mutate(iris3, Sepal.Area = Sepal.Length*Sepal.Width)
```
## 56 Observations and 4 Variables 

# Question 6
```{r}
iris5<-summarize(iris4, meanSepal.Length = mean(Sepal.Length),meanSepal.Width = mean(Sepal.Width), sampleSizeIris4=n())
print(iris5)

```
## 1 Observation and 3 Variables 

# Question 7
```{r}
iris6 <- iris4 %>%
  group_by(Species) %>%
  summarize(meanSepal.Length = mean(Sepal.Length),meanSepal.Width = mean(Sepal.Width), sampleSizeIris4=n())

```
## 2 Observations and 4 Variables

# Question 8
```{r}
iris6b <- iris %>%
  filter(Species %in% c("virginica","versicolor"), Sepal.Length>6, Sepal.Width>2.5) %>%
  select(Species, Sepal.Length, Sepal.Width) %>%
  arrange(by=desc(Sepal.Length)) %>%
  mutate(Sepal.Area = Sepal.Length*Sepal.Width) %>%
  group_by(Species) %>%
  summarize(meanSepal.Length = mean(Sepal.Length),meanSepal.Width = mean(Sepal.Width), sampleSizeIris4=n()) 
```
## Yields the same output

# Question 9
```{r}
iris %>%
  pivot_longer(cols= 1:4,
               names_to = "Measure",
               values_to= "Value")
```


